### 2015-09-02 rscripts/data_review.Rmd
ugh, too many changes

```{r, echo=FALSE, include=FALSE}
req.packages <- c("data.table", "ggplot2", "reshape2", "grid", "scales", "animation", "randomForest")
need.pkgs <- req.packages[!sapply(req.packages, require, character.only = T)]
if (length(need.pkgs) != 0) install.packages(need.pkgs, dependencies = T)
```

@@

```{r, echo=FALSE}
src.dt <- 
  setkey(
    setnames(
      fread("../input/merged.o", header = F, sep=" ", colClasses = list(integer=1.4)),
      c("user_id", "location_id", "login", "logout")
    ),
    login, logout, user_id, location_id
  )
src.dt
```

@@

```{r, echo=FALSE}
source("munging.R")
zeroize(breakoutDays(censor.dt[,
  user_id := .GRP, by=user_id
][,
  location_id := .GRP, by=location_id
], login_day, login_time, logout_day, logout_time, user_id, location_id))
censor.dt
```

>>>>

```{r, echo=FALSE, include=FALSE}
req.packages <- c("data.table", "ggplot2", "reshape2", "grid", "scales", "animation", "randomForest","lubridate")
need.pkgs <- req.packages[!sapply(req.packages, require, character.only = T)]
if (length(need.pkgs) != 0) install.packages(need.pkgs, dependencies = T)
source("munging.R")
```

@@

```{r, echo=FALSE}
src.dt <- loadRaw()
src.dt
```

@@

```{r, echo=FALSE}
zeroize(breakoutDays(censor.dt[,
  user_id := .GRP, by=user_id
][,
  location_id := .GRP, by=location_id
], login_day, login_time, logout_day, logout_time, user_id, location_id))
censor.dt
```

1. consider subset - subset of records? subset of users?  subset of locations?  want predictions about location "class", so subset of locations?  later, when locations classified, subset of users.
2. using subset and some predictor measurements

```{r,echo=F}
## principle components analysis
# weekday_zeroes <- data.table(weekday=factor(c("Sun", "Mon", "Tues", "Wed","Thurs","Fri","Sat"), ordered = T), usage = 0, key = "weekday")
# predictors <- function(SD) with(SD, {
#   duration <- logout - login
#   total_login_time <- sum(duration)
#   usage_by_weekday <- SD[, list(usage = sum(login_day_secs)), keyby=weekday][weekday_zeroes][, list(usage = ifelse(is.na(usage), i.usage, usage)), keyby=weekday]
#   temp <- SD[, sum(logout_day_secs), keyby=weekday][weekday_zeroes][, list(usage = ifelse(is.na(usage), i.usage, usage)), keyby=weekday]
#   usage_by_weekday$usage <- usage_by_weekday$usage + c(temp$usage[7], temp$usage[-7])
#   store <- as.list(usage_by_weekday$usage / total_login_time)
#   names(store) <- usage_by_weekday$weekday
#   reduced_store <- list(weekend=store$Fri+store$Sat, midweek=store$Tues+store$Wed+store$Thurs, weekbegin = store$Sun+store$Mon)
#   user_count <- length(unique(user_id))
#   life <- max(logout) - min(login)
#   return(c(list(
#     log10_lifetime = log10(life),
#     ave_duration = mean(duration),
#     med_duration = median(duration),
#     sd_duration = sd(duration),
#     max_duration = max(duration),
#     total_user_time = total_login_time,
#     log10_user_per_life = log10(user_count/life),
#     log10_time_per_user = log10(total_login_time/user_count),
#     log10_time_per_life = log10(total_login_time/life),
#     log10_unique_users = log10(user_count)
#   ), reduced_store))
# })
# unique_locs <- unique(censor.dt$location_id)
# slice <- sort(sample(unique_locs, size = length(unique_locs)/5, replace = F))
# pca.dt <- censor.dt[, predictors(.SD), by=location_id]
# centered <- pca.dt[,lapply(.SD[,-1,with=F],function(col) col-mean(col)),]
# scaled <- centered[,lapply(.SD, function(col) col/sd(col))]
# prediction.dt <- scaled[slice]
# validation.dt <- scaled[!slice]
# res <- prcomp(prediction.dt[, -1, with=F], scale. = F)
# res$var <- res$sdev^2
# res$pve <- res$var/sum(res$var)
# keep <- res$pve > 0.01 # 1/length(res$pve)
# orig.p <- par(mfrow=c(1,sum(keep)-1))
# for (x in 1:(sum(keep)-1)) biplot(res, choices=c(x,x+1), scale = 0)
# par(orig.p)
# with(res, {
#   plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained", type="b", log="y")
#   abline(h=1/(length(pve)))
#   # abline(h=tail(pve, 1)*2)
#   plot(cumsum(pve), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", ylim=c(0,1), type="b")
#   abline(h=0.9)
#   plot(log10(1-cumsum(pve)), xlab="Principal Component", ylab="Remaining Variance", type="b")
#   abline(a=0, b=log10(1-pve[1]))
#   #plot(pve/(1-c(0,head(cumsum(pve),-1))), xlab="Principal Component", ylab="Proportion of Remaining Variance Explained", ylim=c(0,1), type="b")
# })
# par(orig.p)
# res
# res$x[,]
```

* * *
